{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-09-30 sftrends_manu.Rnw
trends table added

<<trends, fig.height = 6, fig.width = 7, out.width = '0.65\\textwidth', fig.cap = 'Percent changes in nitrogen analytes for (a) annual and (b) seasonal periods in the record.  Changes are based on the difference between the ending and starting estimates for the flow-normalized estimates within each period.  Points are colored for direction (red increasing, green decreasing) and sized for relative magnitude. Station names are shown in the top left panel.'>>=
trnd_map(molabs = NULL, sz_rng = c(2, 8), leg_lab = '(a) Annual % change')
trnd_map(yrlabs = NULL, sz_rng = c(2, 8), stat_labs = F, leg_lab = '(b) Seasonal % change')
@

@@

>>>>

<<trndmap, fig.height = 6, fig.width = 7, out.width = '0.65\\textwidth', fig.cap = 'Percent changes in nitrogen analytes for (a) annual and (b) seasonal (monthly) periods in the record.  Changes are based on the difference between the ending and starting estimates for the flow-normalized estimates within each period.  Points are colored for direction (red increasing, green decreasing) and sized for relative magnitude. Station names are shown in the top left panel.'>>=
trnd_map(molabs = NULL, sz_rng = c(2, 8), leg_lab = '(a) Annual % change')
trnd_map(yrlabs = NULL, sz_rng = c(2, 8), stat_labs = F, leg_lab = '(b) Seasonal % change')
@

@@

% overall trends
<<trnds>>=
data(trnds_chg)
data(trnds_ave)

# percent changes
tab_chg <- trnds_chg

# increase in bold italic
tab_chg[, -c(1:2)] <- round(tab_chg[, -c(1:2)], 1)
tab_chg[, -c(1:2)][sign(tab_chg[, -c(1:2)]) != -1] <- paste0('\\textit{\\textbf{', tab_chg[, -c(1:2)][sign(tab_chg[, -c(1:2)]) != -1], '}}')
tab_chg <- gather(tab_chg, 'var', 'chg', -Site_Code, -resvar) %>% 
  mutate(chg = paste0('\\footnotesize{(', chg, ')}'))

# averages
tab_ave <- gather(trnds_ave, 'var', 'ave', -Site_Code, -resvar) %>% 
  mutate(ave = round(ave, 1))

tab <- left_join(tab_ave, tab_chg, by = c('Site_Code', 'resvar', 'var')) %>% 
  unite(val, ave, chg, sep = ' ') %>% 
  spread(var, val) %>% 
  arrange(resvar, Site_Code)

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and different time periods.  Summaries are averages (mg L$^{-1}$) and percent changes in parentheses (increasing in bold-italic).  Changes are based on the difference between the ending and starting estimates for the estimates within each period.  See \\cref{fig:trndmap} for a summary of spatial trends.'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = c(9, 9, 9),
  cgroup = c('Annual', 'Seasonal'),
  n.cgroup = c(3, 3),
  rowname = stats,
  label = 'tab:trnds'
  )

### 2016-09-30 sftrends_manu.tex
trends table added

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Tue Jul 19 17:50:17 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

@@

>>>>

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Fri Sep 30 16:40:38 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

{\centering \includegraphics[width=0.65\textwidth]{figs/trndmap-1} 
\includegraphics[width=0.65\textwidth]{figs/trndmap-2}

\caption[Percent changes in nitrogen analytes for (a) annual and (b) seasonal (monthly) periods in the record]{Percent changes in nitrogen analytes for (a) annual and (b) seasonal (monthly) periods in the record.  Changes are based on the difference between the ending and starting estimates for the flow-normalized estimates within each period.  Points are colored for direction (red increasing, green decreasing) and sized for relative magnitude. Station names are shown in the top left panel.}\label{fig:trndmap}
\end{figure}

@@

% overall trends
%latex.default(tab[, -c(1, 2)], file = "", rowlabel = "Analyte/Station",     caption = cap.val, caption.loc = "top", rgroup = vars, n.rgroup = c(9,         9, 9), cgroup = c("Annual", "Seasonal"), n.cgroup = c(3,         3), rowname = stats, label = "tab:trnds")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in nitrogen analytes for all stations and different time periods.  Summaries are averages (mg L$^{-1}$) and percent changes in parentheses (increasing in bold-italic).  Changes are based on the difference between the ending and starting estimates for the estimates within each period.  See \cref{fig:trndmap} for a summary of spatial trends.\label{tab:trnds}} 
\begin{center}
\begin{tabular}{llllclll}
\hline\hline
\multicolumn{1}{l}{\bfseries Analyte/Station}&\multicolumn{3}{c}{\bfseries Annual}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Seasonal}\tabularnewline
\cline{2-4} \cline{6-8}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{1976-1988}&\multicolumn{1}{c}{1989-2000}&\multicolumn{1}{c}{2001-2012}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{JFMA}&\multicolumn{1}{c}{MJJA}&\multicolumn{1}{c}{SOND}\tabularnewline
\hline
{\bfseries DIN}&&&&&&&\tabularnewline
~~C10&1.2 \footnotesize{(\textit{\textbf{38.4}})}&1.3 \footnotesize{(\textit{\textbf{16.2}})}&1.2 \footnotesize{(-39.3)}&&1.1 \footnotesize{(-25.7)}&1.1 \footnotesize{(-10.1)}&1.5 \footnotesize{(\textit{\textbf{1.7}})}\tabularnewline
~~C3&0.3 \footnotesize{(\textit{\textbf{33.6}})}&0.4 \footnotesize{(\textit{\textbf{27.4}})}&0.4 \footnotesize{(-12.7)}&&0.4 \footnotesize{(\textit{\textbf{30.7}})}&0.3 \footnotesize{(\textit{\textbf{54.8}})}&0.5 \footnotesize{(\textit{\textbf{42.8}})}\tabularnewline
~~D19&0.4 \footnotesize{(-20.2)}&0.4 \footnotesize{(\textit{\textbf{6.5}})}&0.4 \footnotesize{(-22.6)}&&0.6 \footnotesize{(-28.6)}&0.3 \footnotesize{(-35.1)}&0.4 \footnotesize{(-15.6)}\tabularnewline
~~D26&0.4 \footnotesize{(-2.4)}&0.4 \footnotesize{(\textit{\textbf{10.2}})}&0.4 \footnotesize{(-15.4)}&&0.5 \footnotesize{(-10.9)}&0.3 \footnotesize{(-3.4)}&0.5 \footnotesize{(-2.6)}\tabularnewline
~~D28&0.4 \footnotesize{(-20.5)}&0.4 \footnotesize{(\textit{\textbf{2.7}})}&0.4 \footnotesize{(-32.6)}&&0.6 \footnotesize{(-32.9)}&0.2 \footnotesize{(-53.3)}&0.4 \footnotesize{(-48.8)}\tabularnewline
~~D4&0.3 \footnotesize{(-9.1)}&0.4 \footnotesize{(\textit{\textbf{20.7}})}&0.4 \footnotesize{(-4.3)}&&0.5 \footnotesize{(\textit{\textbf{11.4}})}&0.3 \footnotesize{(\textit{\textbf{4.8}})}&0.4 \footnotesize{(\textit{\textbf{10.7}})}\tabularnewline
~~D6&0.4 \footnotesize{(-9.6)}&0.4 \footnotesize{(\textit{\textbf{21.9}})}&0.5 \footnotesize{(-7.1)}&&0.5 \footnotesize{(-4.7)}&0.4 \footnotesize{(\textit{\textbf{45.6}})}&0.4 \footnotesize{(\textit{\textbf{34.7}})}\tabularnewline
~~D7&0.4 \footnotesize{(-26.4)}&0.4 \footnotesize{(\textit{\textbf{20.4}})}&0.5 \footnotesize{(-2.4)}&&0.5 \footnotesize{(-6.9)}&0.4 \footnotesize{(\textit{\textbf{19}})}&0.4 \footnotesize{(\textit{\textbf{21.4}})}\tabularnewline
~~P8&1.1 \footnotesize{(\textit{\textbf{74.1}})}&1.5 \footnotesize{(\textit{\textbf{37.9}})}&1.6 \footnotesize{(-31)}&&1.4 \footnotesize{(\textit{\textbf{31.6}})}&1 \footnotesize{(\textit{\textbf{52.8}})}&1.8 \footnotesize{(\textit{\textbf{45.4}})}\tabularnewline
\hline
{\bfseries NH$_{4}^{+}$}&&&&&&&\tabularnewline
~~C10&0.1 \footnotesize{(-18.7)}&0.1 \footnotesize{(-58.6)}&0 \footnotesize{(-43.1)}&&0.1 \footnotesize{(-78)}&0 \footnotesize{(-96.2)}&0.1 \footnotesize{(-91.7)}\tabularnewline
~~C3&0.2 \footnotesize{(\textit{\textbf{74.6}})}&0.2 \footnotesize{(\textit{\textbf{43}})}&0.3 \footnotesize{(-18.6)}&&0.2 \footnotesize{(\textit{\textbf{81.8}})}&0.2 \footnotesize{(\textit{\textbf{95.8}})}&0.3 \footnotesize{(\textit{\textbf{74.6}})}\tabularnewline
~~D19&0.1 \footnotesize{(\textit{\textbf{14.5}})}&0.1 \footnotesize{(\textit{\textbf{7.3}})}&0 \footnotesize{(-11.6)}&&0.1 \footnotesize{(-16)}&0 \footnotesize{(-21.8)}&0.1 \footnotesize{(-4.8)}\tabularnewline
~~D26&0.1 \footnotesize{(\textit{\textbf{32}})}&0.1 \footnotesize{(\textit{\textbf{9.3}})}&0.1 \footnotesize{(-17.6)}&&0.1 \footnotesize{(-0.9)}&0.1 \footnotesize{(\textit{\textbf{7.9}})}&0.1 \footnotesize{(\textit{\textbf{18.7}})}\tabularnewline
~~D28&0.1 \footnotesize{(-9.3)}&0 \footnotesize{(-18.2)}&0 \footnotesize{(-12.5)}&&0.1 \footnotesize{(-42.3)}&0 \footnotesize{(-24.7)}&0.1 \footnotesize{(-34.7)}\tabularnewline
~~D4&0.1 \footnotesize{(\textit{\textbf{10.5}})}&0.1 \footnotesize{(\textit{\textbf{47.8}})}&0.1 \footnotesize{(\textit{\textbf{0.2}})}&&0.1 \footnotesize{(\textit{\textbf{66.9}})}&0 \footnotesize{(\textit{\textbf{33.6}})}&0.1 \footnotesize{(\textit{\textbf{68.3}})}\tabularnewline
~~D6&0.1 \footnotesize{(\textit{\textbf{19.9}})}&0.1 \footnotesize{(\textit{\textbf{33.3}})}&0.1 \footnotesize{(-7.3)}&&0.1 \footnotesize{(\textit{\textbf{36}})}&0.1 \footnotesize{(\textit{\textbf{56.4}})}&0.1 \footnotesize{(\textit{\textbf{45.3}})}\tabularnewline
~~D7&0.1 \footnotesize{(-18.1)}&0.1 \footnotesize{(\textit{\textbf{34.4}})}&0.1 \footnotesize{(-0.3)}&&0.1 \footnotesize{(\textit{\textbf{21.2}})}&0 \footnotesize{(-16)}&0.1 \footnotesize{(\textit{\textbf{34}})}\tabularnewline
~~P8&0.2 \footnotesize{(\textit{\textbf{178.6}})}&0.3 \footnotesize{(\textit{\textbf{49.2}})}&0.2 \footnotesize{(-86.1)}&&0.3 \footnotesize{(-52.7)}&0.1 \footnotesize{(-23.9)}&0.3 \footnotesize{(-61.1)}\tabularnewline
\hline
{\bfseries NO$_{2}^{-}$/NO$_{3}^{2-}$}&&&&&&&\tabularnewline
~~C10&1 \footnotesize{(\textit{\textbf{48.1}})}&1.2 \footnotesize{(\textit{\textbf{23}})}&1.1 \footnotesize{(-38.1)}&&1 \footnotesize{(-12.5)}&1 \footnotesize{(-0.2)}&1.4 \footnotesize{(\textit{\textbf{20.4}})}\tabularnewline
~~C3&0.1 \footnotesize{(-4.4)}&0.1 \footnotesize{(\textit{\textbf{11.7}})}&0.2 \footnotesize{(-0.8)}&&0.2 \footnotesize{(-2.3)}&0.1 \footnotesize{(\textit{\textbf{29.4}})}&0.2 \footnotesize{(\textit{\textbf{3.3}})}\tabularnewline
~~D19&0.4 \footnotesize{(-22)}&0.4 \footnotesize{(\textit{\textbf{11.1}})}&0.3 \footnotesize{(-23.7)}&&0.5 \footnotesize{(-28.6)}&0.2 \footnotesize{(-34)}&0.4 \footnotesize{(-16.3)}\tabularnewline
~~D26&0.3 \footnotesize{(-9)}&0.3 \footnotesize{(\textit{\textbf{12.4}})}&0.3 \footnotesize{(-14.8)}&&0.4 \footnotesize{(-13.4)}&0.2 \footnotesize{(-11.8)}&0.3 \footnotesize{(-4.7)}\tabularnewline
~~D28&0.4 \footnotesize{(-29.1)}&0.4 \footnotesize{(\textit{\textbf{7}})}&0.3 \footnotesize{(-34.7)}&&0.5 \footnotesize{(-34.5)}&0.2 \footnotesize{(-59.1)}&0.4 \footnotesize{(-53.8)}\tabularnewline
~~D4&0.3 \footnotesize{(-8.2)}&0.3 \footnotesize{(\textit{\textbf{16.6}})}&0.3 \footnotesize{(-3.7)}&&0.4 \footnotesize{(\textit{\textbf{5.4}})}&0.3 \footnotesize{(\textit{\textbf{6.6}})}&0.3 \footnotesize{(\textit{\textbf{7.3}})}\tabularnewline
~~D6&0.3 \footnotesize{(-9.5)}&0.4 \footnotesize{(\textit{\textbf{16.5}})}&0.4 \footnotesize{(-7.5)}&&0.4 \footnotesize{(-17.9)}&0.3 \footnotesize{(\textit{\textbf{42.6}})}&0.3 \footnotesize{(\textit{\textbf{27.6}})}\tabularnewline
~~D7&0.3 \footnotesize{(-7.5)}&0.4 \footnotesize{(\textit{\textbf{20.5}})}&0.4 \footnotesize{(-7.9)}&&0.4 \footnotesize{(-12.7)}&0.3 \footnotesize{(\textit{\textbf{41.9}})}&0.3 \footnotesize{(\textit{\textbf{24.4}})}\tabularnewline
~~P8&0.9 \footnotesize{(\textit{\textbf{61.3}})}&1.2 \footnotesize{(\textit{\textbf{32.7}})}&1.3 \footnotesize{(-12.5)}&&1.1 \footnotesize{(\textit{\textbf{52.2}})}&0.9 \footnotesize{(\textit{\textbf{60.6}})}&1.4 \footnotesize{(\textit{\textbf{67.3}})}\tabularnewline
\hline
\end{tabular}\end{center}
